### 2009-02-27 doc/ws-programming-tutorial/doc/tex_pesistentservice/peristentservice.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

export ARC\_BARTENDER\_URL

@@

\section{Schlusswort}

Other tutorials are:\\
\\
%Oxana reported the status of OGF tutorials:
\begin{itemize}
 \item first tutorial will be given by Oxana and will deal with how to add the 
resources to Nordugrid, .... (will be related mostly to production ARC)
 \item second tutorial will be given by Gabor and he will show the 
interoperability of ARC (video demo will be shown). A-REX - Unicore 
interoperability.
 \item there will be also a tutorial related to ETICS system where the 
possibility of ETICS system to submit a job to ARC and Unicore will be 
presented. The ETICS team has not contacted so far.
\end{itemize}

>>>>

\task{Some more comments: the <LogLevel> thing is not working at service level, you can just remove those lines - proposed by Zsombor}
export ARC\_BARTENDER\_URL

@@

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_sec_echoservice/sec_echoservice.tex
Started outline for the TLS echo service
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

@@

Currently, the ARC1 comes in severalfour policy decision point implementations:
\begin{itemize}
 \item arc.pdp – Arc PDP\\
The Arc PDP will organize the security attributes into the ARC specific authorization request, call the
policy evaluator to evaluate the request against the policy (which is in ARC specific format) repository,
and get back the evaluation result. See paragraph 3 for detail information about request schema and policy
schema.
\item delegtion.pdp – Delegation PDP\\
The Delegation PDP is basically similar to Arc PDP, except it uses the delegation policy parsed from
remote proxy credential by delegation.collector, and evaluates the request against delegation policy. See
section 6for the design idea and use case of delegation policy in fine-grained identity delegation.
\item simplelist.pdp – Simplelist PDP\\
The Simplelist PDP is a simplest implementation of policy decision point. It will match the identity
extracted from the remote credential (or proxy credential) with local list of permitted identities.
\item pdpservice.invoker – PDP Service Invoker\\
The PDP Service Invoker is a client which can be used to invoke the PDP Service which implements the
same functionality as Arc PDP, except that the evaluation request and response are carried by SOAP
message. The benefit of implementing PDP Service and PDP Service Invoker is that the policy evaluation
engine can be accessed remotely and maintained centrally.
 \item allow.pdp\\
 Does something unknown!
\end{itemize}

@@

>>>>

The TLS echo service, presented in the previous chapter enables the adminstrator the control the access to the grid. Due to a certificate owner may be a member of a large VO it is sometimes recommended to have a more fine grained security within the daemon. 
The goal to be reached in this chapter is to enhance the possibilities to limitate the access to particular users and to narrow their rights on the workstation.\\

The security framework of ARC is implemented based on the same concept of modularity as HED. 
To extend the security of a MCC one only has to modify the service configuration file. This is to be done by including a \textit{SecHandler} element into a component or service which again has to contain a \textit{pdp} element. A small example of a chain component is shown in Listing~\ref{lst:sec_service_component_xml}. The \textit{SecHandler} is able to intervene the message flow
by: manipulating the message content, the attributes or just disrupt the message. 
% Attributes done by: identity.map, delegation.collector
% manipulating content: usernametoken.handler
% disrupt by: arc.authz
\lstsetJUSTXML
\begin{lstlisting}[
label=lst:sec_service_component_xml,float=htb,
caption={[Example of a SecHandler within the component \textit{http.service}.]
         \textbf{Example of a SecHandler within the component \textit{http.service}}}]
<Component name="http.service" id="http">
	<next id="soap">POST</next>
	<SecHandler name="arc.authz" id="auth" event="incoming">
		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
	</SecHandler>
</Component>
\end{lstlisting}
Each MCC or Service usually implement two queues of SecHandlers --- one for incoming messages and one for outgoing called
``incoming'' and ``outgoing'' respectively. All SecHandler components attached to the queue are executed sequentially. If any of them fails, message processing fails as well~\cite{QIANG_2008}.
In order to decide which message shall be modified by the \textit{SecHandler} so called PDP (Policy Decision Point) are used. They come in a bunch of flavours such that the variabilty of security which can be realised is big.\\

In this tutorial only a small set of \textit{SecHandler}s and \textit{PDP} will be introduced. For further reading it is recommended to consult the document ``Security framework of ARC1'' written by Weizhong Qiang and Alexandr Konstantinov. The present chapter is closly related to it.\\

  \begin{table}[htb]
  \centering
  \caption{List of SecHandlers}
  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
	\hline
 	\textbf{identity.map}          & Maps the identity of the user to an identity of the workstation.\\
 	\textbf{arc.authz}             & Restricts the access to the next layer.\\
 	\textbf{delegation.collector}  & Collects the information for the delegation policy (Related to \textit{delegate.pdp}). \\
 	\textbf{usernametoken.handler} & Generates the WS-Security username-token and adds it into the SOAP header of outgoing
					 messages. In case of incoming messages the SecHandler extracts the WS-Security username-token out of it. \task{Figure out what the username-token does...}\\
	\hline
  \end{tabular*}
  \end{table}

  \begin{table}[htb]
  \centering
  \caption{List of PDPs}
  \begin{tabular*}{\textwidth}[t]{lp{12cm}}
	\hline
 	\textbf{allow.pdp}           & bla \\
 	\textbf{simplelist.pdp}      & blub.\\
 	\textbf{arc.pdp}             &\\
 	\textbf{delegate.pdp}        & \task{Explain that well}\\
 	\textbf{pdpservice.invoker}  &\\
	\hline
  \end{tabular*}
  \end{table}

In order to realise security within an MCC 
\begin{verbatim}
SecHandler 
==========
	PDP
	===	
		elements/attributes required
		============================
identity.map						--- maps the identity of users to identity within the local workstation
	allow.pdp					--- means concering all identities
							  (I suppose it can be used in every MCC)
		<LocalName>nobody</LocalName>		  (Due to the SecHandler within the PDP the local name has to be spechified)
		???LocalList???
		???LocalSimplePool???
	simplelist.pdp					--- maps only persons listed within the file
								(TLS identity is used here, thus it can be used only after the TLS)
		<PDP name="simplelist.pdp" location="./clientlist.txt"/>
	arc.pdp						--- Complexer policy to permit or deny someone. 
		<PolicyStore>					(Is it possible to add parameters for the SOAP service here?)
		<Location type="file">policy.xml</Location>
		</PolicyStore>
	[...]

arc.authz					--- restricts access to next layer
	arc.pdp					--- using policy.xml
	simplelist.pdp				--- using a file with listed identities
	delegate.pdp				--- like arc.pdp but poicy is parsed from a remote proxy
	pdpservice.invoker			--- Is a client which can be used to invoke the PDP Service which 
							implements the same functionality as Arc PDP, except that the
							evaluation request and response are carried by SOAP
							message. The benefit of implementing PDP Service and PDP Service 
							Invoker is that the policy evaluation engine can be accessed remotely
							and maintained centrally.
							- wasn't that also possible with the delegate.pdp??

delegation.collector 				--- Delegation SecHandler
						--- is responsible for collecting the delegation policy information and 
						putting this policy into message’s security attribute for the 
						usage of other components, such as delegation.pdp
usernametoken.handler				--- The task is to generate the WS-Security Username-Token and add it into
						header of SOAP message which is the payload of outgoing message. 
						It can also extract the WS-Security Username-Token from the header of 
						SOAP message which is the payload of incoming message.
						why should one want to do that???

\end{verbatim}

- have fine grained but complex authorization system (policy evaluator) which uses information stored on the server (arc.pdp) or an an proxy (delegation.pdp) or which is evaluated using a SOAP based service such that the request can be evaluated remotly (pdpservce.invoker). 
- have a simple list which matches identities (simplelist.pdp)
- map users to an identity within the local workstation

have a smaller set of users,
have a policy which may map certain users to certain settings

For HED is build up to  % entsrprechend einer Mod Structur
be modular, the security can easily be extend to our given example.
Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler. The SecHandler components provide a method for processing messages traveling through Message Chains of the HED. 
Each MCC or Service usually implement two queues of SecHandlers – one for incoming messages and one for outgoing called “incoming” and “outgoing” respectively. All SecHandler components attached to the queue are executed sequentially. 
If any of them fails, message processing fails as well. 
Each SecHandler is configured inside the arched configuration file used for configuring whole chain of MCCs 
Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP). PDP can process ARC specific Request and Policy documents which are as well written in XML format~\cite{QIANG_2008}..

@@

@@

\section{Other PDPs}

In order the adminisitrator wants to maintain the grid access using one workstation the delegation.pdp can be used.
The client requests access to a service. In case the request is valid a certificate will be deployed on the desired server such that it can be accessed now \task{Am I right?} \cite{QIANG_2008}\\

### 2009-02-27 doc/ws-programming-tutorial/doc/tex_tls_echoservice/tls_echoservice.tex
Express checkout... Finished the basic TLS chapter content and started writting about the Sec echo client...

\clearpage
\section{Service}

In order to create a secure bilateral connection we hence need two certificates: a client certificate and a service certificate.
The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key. 
Altogether six files are needed:
\begin{tabular}{l@{-}p{9cm}}
clientCA   & The CA which guarantees for the identity of the client \\
clientCERT & The certificate of the client. \\
clientKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
\\
serverCA   & The CA which guarantees for the identity of the server \\
serverCERT & The certificate of the server. \\
serverKEY  & Secret key of the client to read messages, encrypted by the public key in the certificate! \\
\end{tabular}

The server needs: its  serverCERT and serverKEY and some kind of root CA knowing the client.
Client needs its clientCERT and clientKEY and some kind of root CA knowing the server.

\task{RENAMING into~ARC service configuration file}
\task{echo service client configuration file}

%The first two goals can be reached by encrypting the channel/connection between client and server. An approach 
%sich bewähren / behaupten
%which stand the test of time is to introduce a TLS (Transport Layer Security) also known as SSL (Secure Sockets Layer).
%Authenticity, Confidentiality, Integrity
%The client wants to reassure himself to send his vulnerable data only to the server who can prove its correct identity.
%Furthermore a prove of authenticity is needed. % Sich versichern, dass das Gegenüber der ist, für welchen er sich ausgibt
%It .... does fancy stuff ...
%
%
%Authorization:  The server on the other side takes an interest in limitate its ressources to a close circle of users. Above all no %foreign clients shall be able to use the services.\\

One can add the TLS in client and server by simply extending the arched configuration file and a small modification of the client.

WARNING: If the program fails, check if one of your certificates has expired.
If so, create new one using the script files in the directory 'certFactory'.

Chalked up to the modularity of HED, the security can easily be extend to our given example.
Each Message Chain Component (MCC) or service has a common interface for implementing various pluggable components (plug-ins) called SecHandler.
The SecHandler components provide a method for processing messages traveling through Message Chains of the HED.

Each MCC or Service usually implements two queues of SecHandlers – one for incoming messages and one for outgoing, which are called “incoming” and “outgoing”, respectively.
All SecHandler components attached to the queue are executed sequentially. 
If any of them fails, message processing fails as well. 
Each SecHandler is configured inside the arched configuration file that is used for configuring whole chain of MCCs 
Some of the currently implemented SecHandler components make use of pluggable and configurable sub-modules called Policy Decision Point (PDP).
PDPs can process ARC specific Request and Policy documents which are also written in XML format~\cite{QIANG_2008}..

4 types:
                <KeyPath>./clientKey.pem</KeyPath>
                <CertificatePath>./clientCert.pem</CertificatePath>
                <CACertificatePath>./serviceCA.pem</CACertificatePath>

\lstsetCPP
\lstinputlisting
	[
	label=lst:tls_echo_service_cpp,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/tlsechoservice/tlsechoservice.cpp}

\lstsetARCHEDXML
\lstinputlisting
	[
	label=lst:tls_echo_arched_xml,float=htb,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/tlsechoservice/arched_tls_echoservice.xml}

\lstsetKSH
\begin{lstlisting}[
label=lst:tls_echo_arched_invoke,float=htb,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ rm -f /var/log/arched.log
$ arched -c arched_echoservice.xml  && echo jo ||echo n
$ tail -n100 -f /var/log/arched.log
$ killall arched
\end{lstlisting}

@@

\lstsetCPP
\lstinputlisting
	[
	label=lst:tls_echo_client_cpp,
	caption={[HEC configuration file]
	\textbf{HEC configuration file\textcolor{white}{hmf}}}
	]
{../src/clients/tlsechoclient/tlsechoclient.cpp}

Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:

%[caption={[]WegDamit},language=XML,basicstyle=\scriptsize,breaklines=true,label=lst:request]

Zertifikate
Zustände (Nutzer wiedererkennen, arbeit aufnehmen)

\lstsetKSH
\begin{lstlisting}[
label=lst:tls_echo_client_invoke,float=htb,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
$ ./echoclient
Tue Feb 17 17:04:08 2009
\end{lstlisting}

\lstsetJUSTXML
\begin{lstlisting}[
label=lst:tls_echo_request_XML,float=htb,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <tlsecho:tlsechoRequest>
      <tlsecho:say operation="ordinary">text_to_be_transmitted</tlsecho:say>
    </tlsecho:tlsechoRequest>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}

\lstsetJUSTXML
\begin{lstlisting}[
label=lst:tls_echo_response_XML,float=htb,
caption={[Transformation in eine uniforme konzentrische Verteilung.]
         \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
<soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
  <soap-env:Body>
    <tlsecho:tlsechoResponse>
      <tlsecho:hear>[ text_to_be_transmitted ]</tlsecho:hear>
    </tlsecho:tlsechoResponse>
  </soap-env:Body>
</soap-env:Envelope>
\end{lstlisting}

>>>>

\clearpage
\section{Service}
\task{RENAMING HED configuration file into~ARC service configuration file - proposed by Marek}
\task{(echo service) client configuration file - proposed by Marek}

In order to create a secure bilateral connection we hence need two certificates: a client certificate and a service certificate.
The CA for the certificate has to be pre-configured on the counterpart and furthermore, the owner of the certificate has to hold the fitting private key.
For the given example X.509-certificates will be used which are standardised by the ITU-T.
Altogether six files are needed:
\begin{tabular}{l@{~---~}p{13cm}}
\textbf{clientCA}   & The CA which guarantees for the identity of the client \\
\textbf{clientCERT} & The certificate of the client. \\
\textbf{clientKEY}  & The private key of the client (should be kept in safe custody).\\
\\
\textbf{serverCA}   & The CA which guarantees for the identity of the server \\
\textbf{serverCERT} & The certificate of the server. \\
\textbf{serverKEY}  & The private key of the server (should be kept in safe custody).\\
\end{tabular}
\forcelinebreak
As one can see, PEM has been choosen for the file format.\\

To get this service running, the six files have to be distributed. The server should hold the files: serverCERT.pem, serverKEY.pem and clientCA.pem. The client on the other side should own the files: clientCERT.pem, clientKEY.pem and serverCA.pem.
Within the source code directory certificates with corresponding names are prepared. Due to certificates have a limited period of validity, these certificates may no longer be accepted by the communication parter. In order to create a first basic set of certificates signed by a self created CA, the shell scripts \textit{createMasterCA.sh} and \textit{createSlaveCert.sh} to be found in the directory \textit{src/services/tlsechoservice/certFactory} may be used. The script \textit{createMasterCA.sh} will create a new CA and overwrite the old on which was stored in the directory. The script \textit{createSlaveCert.sh} create an arbitary amount of certificates signed by the CA which has been create by the other script.\\
%WARNING: If the programm fails, check if one of your certifacte is expired. If so, create new one using the script files in the directory 'certFactory'

The TLS echo service implementation is almost the same as the one presented in the chapter before. Merely the Class name and the namespace of the service has changed. The only modification has to be done within the ARC service configuration file which is displayed in Listing~\ref{lst:tls_echo_arched_xml}.

\lstsetARCHEDXML
\lstinputlisting
	[
	label=lst:tls_echo_arched_xml,float=htb,
	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
	]
{../src/services/tlsechoservice/arched_tls_echoservice.xml}
The first change is to be found in lin~\ref{lst_code:tls_echo_arched_xml_plugin}. As to be seen, another plugin called \textit{mcctls} will be loaded by the \textit{ModuleManager}. The second change is the new element which has been placed between the TCP and the HTTP component, see line~\ref{lst_code:tls_echo_arched_xml_component}. Within the element the files are declared using the elements
\begin{verbatim}
<KeyPath>./clientKey.pem</KeyPath>
<CertificatePath>./clientCert.pem</CertificatePath>
<CACertificatePath>./serviceCA.pem</CACertificatePath>
<CACertificatesDir>/etc/grid-security/certificates</CACertificatesDir>
\end{verbatim}
or
\begin{verbatim}
<ClientSSLConfig FromFile='filename'/>
\end{verbatim}
\task{You can put the ClientSSLConfig into a separate file, then use <ClientSSLConfig FromFile='filename'/> in each service, and then you don't need to change everywhere - proposed by Zsombor}

%\lstsetCPP
%\lstinputlisting
%	[
%	label=lst:tls_echo_service_cpp,
%	caption={[HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml]
%	\textbf{HED configuration file for the Arc intern echo service. Filename: arcecho\_no\_ssl.xml\textcolor{white}{hmf}}}
%	]
%{../src/services/tlsechoservice/tlsechoservice.cpp}
%
% Lassen wir mal weg....
%
%
% \lstsetKSH
% \begin{lstlisting}[
% label=lst:tls_echo_arched_invoke,float=htb,
% caption={[Transformation in eine uniforme konzentrische Verteilung.]
%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
% $ rm -f /var/log/arched.log
% $ arched -c arched_echoservice.xml  && echo jo ||echo n
% $ tail -n100 -f /var/log/arched.log
% $ killall arched
% \end{lstlisting}
% 
%        ...das auch.....

@@

In this section a new way to implement the client will be introduced. As already mentioned in section~\ref{sec:timeservice_client}, the protocol stack of the client is build up in almost the same way like the service. Indeed, the class~\textit{ClientSOAP} composes the stack with the same MCC used for the service. Once this is known, its self-evident to create the client in a simlar manner like the service with an own configuration file.
Listing~\ref{lst:tls_echo_client_cpp} shows the C++ source code of the client implementation. The main difference between the simple echo client presented in the chapter before is that now the information about the client protocal stack is encapsulated within a configuration file.
\lstsetCPP
\lstinputlisting
	[
	label=lst:tls_echo_client_cpp,
	caption={[HEC configuration file]
	\textbf{HEC configuration file\textcolor{white}{hmf}}}
	]
{../src/clients/tlsechoclient/tlsechoclient.cpp}
The file containing the client configuration is specified within the command line. 
Its content will be loaded in the code fragment starting in line~\ref{lst_code:tls_echo_client_cpp_getfilecontent}. 
Later, in line~\ref{lst_code:tls_echo_client_cpp_loadconfiguration}, the configuration will be parsed into XML and than transfered into a chain of component which is wrapped by the object \textit{loader}. In order to process a message, an entry point within the chain has to be disclosed. For we are implementing a SOAP client, we are looking for the SOAP component. 
To get the proper component an operator of MCCLoader is used in line~\ref{lst_code:tls_echo_client_cpp_cliententry}.
The previously used class \textit{ClientSOAP} has eased some workload which now has to be done within the method.
The payload now have to be wrapped by the message class which have to be initialised with an \textit{MessageAttribute} object and a \textit{MessageContext} object. After the payload has been wrapped within the message it gets transmitted in almost the same way like before, see lines followed by line~\ref{lst_code:tls_echo_client_cpp_createandprocess}. The rest of the source code is identical to the echo client.\\

% Folgender XML Aufruf und Antwort soll \textit{automatisch} generiert werden:

% 
% \lstsetJUSTXML
% \begin{lstlisting}[
% label=lst:tls_echo_request_XML,float=htb,
% caption={[Transformation in eine uniforme konzentrische Verteilung.]
%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
% <soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
%   <soap-env:Body>
%     <tlsecho:tlsechoRequest>
%       <tlsecho:say operation="ordinary">text_to_be_transmitted</tlsecho:say>
%     </tlsecho:tlsechoRequest>
%   </soap-env:Body>
% </soap-env:Envelope>
% \end{lstlisting}
% 
% 
% 
% 
% 
% 
% \lstsetJUSTXML
% \begin{lstlisting}[
% label=lst:tls_echo_response_XML,float=htb,
% caption={[Transformation in eine uniforme konzentrische Verteilung.]
%          \textbf{Transformation in eine uniforme konzentrische Verteilung.\textcolor{white}{hmf}}}]
% <soap-env:Envelope xmlns:tlsecho="urn:tlsecho" xmlns:soap-enc="http://schemas.xmlsoap.org/soap/encoding/" xmlns:soap-env="http://schemas.xmlsoap.org/soap/envelope/" xmlns:xsd="http://www.w3.org/2001/XMLSchema" xmlns:xsi="http://www.w3.org/2001/XMLSchema-instance">
%   <soap-env:Body>
%     <tlsecho:tlsechoResponse>
%       <tlsecho:hear>[ text_to_be_transmitted ]</tlsecho:hear>
%     </tlsecho:tlsechoResponse>
%   </soap-env:Body>
% </soap-env:Envelope>
% \end{lstlisting}
%
